A Linear Solution of the Linearity Challenge

### 2025-12-28 paper/conclusion.tex
tuning related and repo references

The simple formalisation of \LCC is a good starting point for further
developments. We have already extended \LCC with support for infinite types and
recursive processes (the source code of this extension is publicly available in
the same repository as the one that contains the formalisation of \LCC). In the
future, it would be interesting to formalise the strong normalisation property
of \LCC as a consequence of cut elimination of classical linear logic.

>>>>

The simple formalisation of \LCC is a good starting point for further
developments. We have already extended \LCC with support for infinite types and
recursive processes (the source code of this extension is publicly available in
the same repository as the one that contains the formalisation of \LCC~\cite{PadovaniRaffaelli25}). In the
future, it would be interesting to formalise the strong normalisation property
of \LCC as a consequence of cut elimination of classical linear logic.

### 2025-12-28 paper/introduction.tex
tuning related and repo references

In this paper we approach the linearity challenge from a different angle:
\emph{instead of proposing new techniques that make it easy to formalise the
calculus in the challenge, we propose a (relatively) new calculus that is easy
to formalise with the existing techniques}.
%
More specifically, we describe the Linear Calculus of Continuations (\LCC) whose
type system coincides with the proof system of classical linear logic and that
features \emph{linear channels} instead of sessions. Linear channels are meant
to be used for a \emph{single} communication. Structured conversations can be
modeled in \LCC using \emph{explicit continuations} as in the well-known
encoding of (binary) sessions into the linear
$\pi$-calculus~\citep{KobayashiPierceTurner99,Kobayashi02b,DardhaGiachinoSangiorgi17}:
each message exchanged on a linear channel may include one or two fresh channels
-- the \emph{continuations} -- on which the rest of the conversation takes
place.
%
Overall, \LCC is nothing but a lower-level, linear version of
\CP~\cite{Wadler14}, Wadler's calculus of sessions based on classical linear
logic, such that sessions are encoded instead of being featured natively.

@@

The formalisation has been checked with Agda 2.8.0 and the code is available on
GitHub at \url{https://github.com/boystrange/LinearityChallenge/tree/main/lcp}.

>>>>

In this paper we approach the linearity challenge from a different angle:
\emph{instead of proposing new techniques that make it easy to formalise the
calculus in the challenge, we propose a (relatively) new calculus that is easy
to formalise with the existing techniques}.
%
More specifically, we describe the Linear Calculus of Continuations (\LCC) whose
type system coincides with the proof system of classical linear logic and that
features \emph{linear channels} instead of sessions. Unlike session endpoints,
which can be used multiple times albeit sequentially, linear channels are meant
to be used \emph{once}. \LCC retains the expressiveness of modeling structured
conversations using \emph{explicit continuations}, just like in the well-known
encoding of (binary) sessions into the linear
$\pi$-calculus~\citep{KobayashiPierceTurner99,Kobayashi02b,DardhaGiachinoSangiorgi17}:
each message exchanged on a linear channel may include one or two fresh channels
-- the \emph{continuations} -- on which the rest of the conversation takes
place.
%
Overall, \LCC is nothing but a low-level version of \CP~\cite{Wadler14} --
Wadler's calculus of sessions based on classical linear logic -- such that
sessions are encoded instead of being featured natively.

@@

The formalisation has been checked with Agda 2.8.0 and the code is available on
in a public repository on GitHub~\cite{PadovaniRaffaelli25}.

### 2025-12-28 paper/related.tex
tuning related and repo references

\citet{RouvoetPoulsenKrebbersVisser20} present a library of abstractions
inspired to separation logic aiding the formalisation of interpreters for
languages with linear resources. One of the presented case studies is the
formalisation of a fragment of \GV~\citep{GayVasconcelos10,Wadler14}. The
separating conjunction introduced in \Cref{sec:context-agda} and the typing of
the constructors for the representation of processes in \Cref{sec:process-agda}
have been adapted from this work.

>>>>

\citet{RouvoetPoulsenKrebbersVisser20} present a library of abstractions
inspired to separation logic aiding the formalisation of interpreters for
languages with linear resources. One of the presented case studies is the
formalisation of a fragment of \GV~\citep{GayVasconcelos10,Wadler14}. 
%
Unlike the other formalisations we are discussing,
\citet{RouvoetPoulsenKrebbersVisser20} do not define a small-step semantics for
\GV which explains (in part) the small size of their development. However, since
their formalisation of \GV is intrinsically typed, the interpreter preserves
typing and therefore proves a form of subject reduction property.
%
The separating conjunction introduced in \Cref{sec:context-agda} and the typing
of the constructors for the representation of processes in
\Cref{sec:process-agda} have been adapted from this work of
\citet{RouvoetPoulsenKrebbersVisser20}.
